\section{Discussion}

%Scalability?
The current system is a proof of concept that aims to prove that rule patterns
can be extracted from stories in such a way that both similarities in
precondition and effect of the action can be observed. The model also
facilitates creativity, since variations of the precondition and effect are
still possible, and for each possibility it is possible to obtain probabilities.
This captures the creativity of the way a rule was used in the context of a
particular story, but it also enables a program to choose an action based on the
probability of the rule. For variation from the norm, it might choose an action
with low probability, but if
this gets out of hand, it could also choose an action with a high probability to
attempt to make the story more familiar to the reciever.

As discussed in section 4.3.3, a generation program could also analyze character co-occurence
(which characters occur regularly in the same actions). Also, since the
occurence of a character in a certain action is stored, it is possible to
analyze which characters often appear in the same positions.
For example, a king
and a sultan might fulfill the same function of a authority figure in a story,
and therefore appear in many of the same actions.

Another way the system facilitates creativity is that the rule objects are
defined over story actions, instead of longer series of actions as discussed in
the related work section. This ensures more freedom, since actions are not
dependent on each other directly, but only indirectly, since certain
preconditions have to be satisfied.

Because of this, the current
system is limited at the moment by a relatively simple tree grammar to extract
the actions from sentences in natural language. Because of this the sentences
were limited in complexity. This can be solved in several ways. One way would be
to expand the grammar rules of the tree, but the more rules the tree has, the
more parse ambiguity could become a problem.
%How general?
%RDF?
%Structures that capture creativity vs. creativity in generation
%What pieces are useful for eachother?
%Something about how to model the discourse part of a narrative (Chatman)
