\section{Evaluation}

The system has been tested on two stories: `The Frog Prince' (\cite{frogprince}) and
`Caliph Stork' (\cite{stork}). The stories were manually
annotated so they conformed to the state-action model proposed by this thesis.
In section 5.1 the stories as annotated will be presented and the similarities
and differences between them will be discussed, and in section 5.2 an example
involving magical transformations will be shown

\subsection{Dataset}

`The Frog Prince' and `Caliph Stork' were chosen since they are both from the same genre (fairy tales)
and both stories contain a common story element (human to animal transformations). In `The
Frog Prince' the titular prince starts as a frog and eventually becomes a human
prince, in `Caliph Stork' the caliph and his vizier start as humans, are changed
into storks with a magic powder and both have to find a way to get back to human
form. Since the stories have this common element, the program has to be able to
handle this appropriately (extract similar rules for each occurence).

The actions of `The Frog Prince' were as follows:
\begin{verbatim}
the frog promises the ball to the princess
the princess promises companionship to the frog
the frog gets the ball
the princess goes home. the princess has dinner with the king
the frog asks the princess for promise:2. the princess refuses promise:2
the king enforces the promise:2
the princess uses companionship
the frog transforms into a human
the frog marries the princess
\end{verbatim}

The actions of `Caliph Stork' were the following:
\begin{verbatim}
the caliph buys the magic_powder
caliph learns about magic_powder
caliph uses magic_powder. caliph says magic_word. 
caliph uses magic_powder. caliph says magic_word. the caliph transforms into an animal
vizier uses magic_powder. vizier says magic_word. the vizier transforms into an animal
the caliph listens to an animal
the vizier listens to an animal
the caliph laughs
the vizier laughs
mirza becomes the lord_of_baghdad
the caliph meets the owl
the vizier meets the owl
SUBSTORY_BEGIN
	king_of_the_indies denies kaschnur
	the owl transforms into an animal
SUBSTORY_END
caliph learns the  magic_word
vizier learns the magic_word
caliph transforms into a human
vizier transforms into a human
caliph marries the owl
the owl transforms into a human
caliph kills kaschnur
mirza transforms into an animal
\end{verbatim}

Even though both stories contain transformations, they vary in complexity with
respect to the rules. In
Caliph Stork, the transformation was done with a magic powder, the character had to
remember a magic word, and if the character
laughed while in animal form, they forgot the magic word and could not change back. 
This is explicitly told to the caliph, and by extension, to the reader. 
However, in `The Frog Prince'
there are also certain prerequisites for the frog to become a prince again, but
these are never explicitly stated (did the frog have to be kissed by a princess,
or would any person, male or female, do?). Because of this, no explicit
annotation has been done about the rules of the magic transformation in the frog
prince, which results in a much simpler set of prerequisites compared to Caliph
Stork.

\subsection{Extraction of transformation}

To evaluate if the program was able to capture this information, it was given
both stories with the states annotated in graph form and the actions in
simplified natural language. The only other information it was given was the
grammar to parse the action sentences.

Both stories contain 8 occurences of a transformation. There is one animal to human
transformation in `The Frog Prince', and there are 4 transformations from human 
to animal and 3 transformations from animal to human in `Caliph Stork'.
The example from `The Frog Prince' (where the story event is `the frog
transforms into a human') looks as follows:

\begin{verbatim}
OBJ  = `human'
SUB  = `frog'
PREC = [
    (`SUB', `SUB_race:animal'),
    (`SUB_abilities:ball', `SUB'), 
    (`SUB', `SUB_abilities:ball'),
    (`SUB_race:animal', `SUB')
    ]
EFF = {
    `edges_gone': [
        (`SUB', `SUB_race:animal')
        ],
    `edges_appeared': [
        (`SUB_race:OBJ', `SUB')
        ]
    }
\end{verbatim}

A similar situation, where the caliph in `Caliph Stork' transforms from animal 
to human form is given here (The story event being `caliph transforms into a
human'):

\begin{verbatim}
OBJ  = `human'
SUB  = `caliph'
PREC = [
    (`SUB_abilities:magic_powder', `SUB'),
    ('enemy:1', `kaschnur'),
    (`SUB_abilities:magic_powder',
    `rule:know(x,magic_word)&&speak(x,magic_word)&&animal(x)->OBJ(x)'),
    (`SUB_abilities:magic_powder',
    `rule:know(x,magic_word)&&speak(x,magic_word)&&OBJ(x)->animal(x)'),
    (`SUB_abilities:magic_powder', `rule:animal(x)&&laugh(x)->not(know(x, magic_word))'),
    (`SUB_abilities:wealth', `SUB'), 
    (`SUB', 'enemy:1'), 
    (`SUB', `SUB_race:animal'),
    (`SUB', `SUB_abilities:wealth'),
    (`SUB', `SUB_abilities:know(SUB, magic_word)'),
    ('enemy:1', `SUB'), 
    (`SUB_abilities:know(SUB, magic_word)', `SUB'),
    (`SUB', `SUB_abilities:magic_powder'), 
    (`SUB_race:animal', `SUB')
    ]
EFF = {
    `edges_gone': [
        (`SUB', `SUB_race:animal')
        ],
    `edges_appeared': [
        (`SUB', 'enemy:1'),
        (`SUB_abilities:wealth', `SUB'),
        (`kaschnur', 'enemy:1'),
        (`SUB_race:OBJ', `SUB')
        ]
    }
\end{verbatim}

Another example of the same rule is from earlier in the story, when the caliph
transforms from human to animal by using a magical powder:

\begin{verbatim}
OBJ    = `animal'
SUB    = `caliph'
PREC   = [
    (`SUB_abilities:know(SUB, magic_word)', `SUB'),
    (`SUB_abilities:magic_powder', `SUB'),
    (`SUB', `SUB_race:human'),
    (`SUB_abilities:magic_powder', 
    `rule:know(x,magic_word)&&speak(x,magic_word)&&OBJ(x)->human(x)'), 
    (`SUB_race:human', `SUB'),
    (`SUB_abilities:lord_of_baghdad', `SUB'),
    (`SUB_needs:speak_with_animals', `SUB'),
    (`SUB_abilities:wealth', `SUB'),
    (`SUB', `SUB_abilities:wealth'),
    (`SUB', `SUB_abilities:know(SUB, magic_word)'),
    (`SUB_abilities:magic_powder',
    `rule:know(x,magic_word)&&speak(x,magic_word)&&human(x)->OBJ(x)'),
    (`SUB', `SUB_needs:speak_with_animals'),
    (`SUB', `SUB_abilities:magic_powder'),
    (`SUB', `SUB_abilities:lord_of_baghdad'),
    (`SUB_abilities:magic_powder', 
    `rule:OBJ(x)&&laugh(x)->not(know(x, magic_word))')
    ]
effects: {
    `edges_gone': [
        (`SUB', `SUB_race:human')
        ],
    `edges_appeared': [
        (`SUB_abilities:wealth', `SUB'),
        (`SUB_race:OBJ', `SUB')
        ]
    }
\end{verbatim}

There are similarities between all three occurences. They all share one of the
effects, namely that the race of the subject becomes the object
(\texttt{`SUB\_race:OBJ', `SUB')}.
This is a good property, as it facilitates this rule to be applied generally for all
transformations in future stories, even if the object that the person has been
transformed in has not been observed in another story.

Another common property of these rules is that the subject loses their own race
(this is denoted by \texttt{(`SUB', `SUB\_race:animal')} being one of the
effects of the transformation in the first two examples and 
\texttt{(`SUB', `SUB\_race:human')} being the effect in the third example). However, this similarity is
less pronounced than the previous one, since the race the character started out
with was not specified in the action. The first example was obtained from the
story event \texttt{`the frog transforms into a human.'}. If the story event had
consistently been in the form of \texttt{`the
frog transforms from an animal into a human.'}, then the resulting rule that
would have been learned would have been more general like the previous rule, 
since the race that the frog used to have (animal) would have its own variable,
making it more similar to a transformation in the opposite direction (for
example \texttt{the caliph transforms into an animal}).
%TODO try again, apparently

As said earlier in this section, the transformations in `Caliph Stork' were different from the one in `The Frog
Prince', since in `Caliph Stork' there was a magical powder with its associated
rules (the character using the powder had to speak a magical word to transform, and would forget the
word if they laughed in animal form), while in `The Frog Prince' there was no
explicit rule governing the magic. The rules of this magic powder can be seen in
the preconditions, since they were also represented as rule nodes connected to
the characters who used the powder to transform. For `The Frog Prince' there was no
specific rule, so there are no additional rules added to the precondition. In a
generation scenario, the program could choose between these examples, either
generating a scenario with complex magical rules (as in `Caliph Stork') or a 
scenario without these rules (like in `The Frog Prince').
This highlights that
though the knowledge base is able to recognize regularities in the data (like
the changing of the subject's race), it is also able to represent diversity in
its rule system (a generation program has access to diverse examples of
variations within the same rule).

Of course, the evaluation has been done on just two stories, since the
annotation for now is a time-intensive process. Because of this no actual story
generation has been done, but this thesis should be seen as more of a proof of
concept: patterns can be recognized, and if more annotated stories were
available (either through annotation or, more feasibly, through more
sophisticated natural language processing), the program could learn more interesting patterns. When it is
possible to automatically annotate the stories, it should be more feasible to
generate actually interesting stories, since the program will be able to learn a
larger amount and variety of rules.
